---
title: "Github1"
output: html_document
date: "2023-05-04"
---

#End to end data analysis 
#Data obtained from QFISH database 


```{r}
library(tidyverse)
```


```{r}
iris # look at iris
str(iris) # check it out - what type is it? how many rows?

as_tibble(iris)
#> # A tibble: 150 × 5
#>   Sepal.Length Sepal.Width Petal.Length Petal.Width Species
#>          <dbl>       <dbl>        <dbl>       <dbl> <fct>  
#> 1          5.1         3.5          1.4         0.2 setosa 
#> 2          5.9         3            1.4         0.2 setosa 
#> 3          5.7         3.2          1.3         0.2 setosa 
#> 4          5.6         3.1          1.5         0.2 setosa 
#> 5          5           3.6          1.4         0.2 setosa 
#> 6          5.4         3.9          1.7         0.4 setosa 
#> # ... with 144 more rows


```

```{r}
tibble(
  x = 1:5, 
  y = 1, 
  z = x ^ 2 + y) # call new variables to produce new column values!


#> # A tibble: 5 × 3
#>       x     y     z
#>   <int> <dbl> <dbl>
#> 1     1     1     2
#> 2     2     1     5
#> 3     3     1    10
#> 4     4     1    17
#> 5     5     1    26


```

```{r}
data.frame(c(
     x = 1:5, 
     y = 1, 
     z = x ^ 2 + y
))
Error in data.frame(c(x = 1:5, y = 1, z = x^2 + y)) : 
  object 'x' not found


```
```{r}
tb <- tibble(
  `:)` = "smile", 
  ` ` = "space",
  `2000` = "number"
)
tb
#> # A tibble: 1 × 3
#>   `:)`  ` `   `2000`
#>   <chr> <chr> <chr> 
#> 1 smile space number
```
```{r}
tribble(
  ~x, ~y, ~z,
  #--|--|----
  "a", 2, 3.6,
  "b", 1, 8.5
)
#> # A tibble: 2 × 3
#>   x         y     z
#>   <chr> <dbl> <dbl>
#> 1 a         2   3.6
#> 2 b         1   8.5

```

```{r}
as_tibble(iris)
```

```{r}
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)
#> # A tibble: 1,000 × 5
#>   a                   b              c     d e    
#>   <dttm>              <date>     <int> <dbl> <chr>
#> 1 2022-11-18 23:15:18 2022-11-25     1 0.368 n    
#> 2 2022-11-19 17:20:28 2022-11-30     2 0.612 l    
#> 3 2022-11-19 11:44:07 2022-12-10     3 0.415 p    
#> 4 2022-11-19 01:05:24 2022-12-09     4 0.212 m    
#> 5 2022-11-18 21:29:41 2022-12-06     5 0.733 i    
#> 6 2022-11-19 08:30:38 2022-12-02     6 0.460 n    
#> # ... with 994 more rows


```

```{r}
install.packages("nycflights13")
library(nycflights13)
nycflights13::flights %>% 
  print(n = 10, width = Inf)

```



```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

# Extract by name
df$x
#> [1] 0.73296674 0.23436542 0.66035540 0.03285612 0.46049161
df[["x"]]
#> [1] 0.73296674 0.23436542 0.66035540 0.03285612 0.46049161

# Extract by row position
df[[1]]
#> [1] 0.73296674 0.23436542 0.66035540 0.03285612 0.46049161

# Extract by exact position
df[[2,2]]
#[1] -1.751506

```

```{r}

df %>% .$x
#> [1] 0.73296674 0.23436542 0.66035540 0.03285612 0.46049161
df %>% .[["x"]]
#> [1] 0.73296674 0.23436542 0.66035540 0.03285612 0.46049161


```


```{r}
df <- tibble(
    xxx = runif(5),
    y = rnorm(5)
)
df$xx
NULL
Warning message:
Unknown or uninitialised column: `xx`. 

```
```{r}
df <- data.frame(abc = 1, xyz = "a")
df
df$x #call by name
df[, "xyz"] #call by exact position

```

#Importing data 

#> Rows: 1192 Columns: 6
#> ── Column specification ────────────────────────────────────────────────────────
#> Delimiter: ","
#> chr (2): sex, race
#> dbl (4): earn, height, ed, age
#> 
#> ℹ Use `spec()` to retrieve the full column specification for this data.
#> ℹ Specify the column types or set `show_col_types = FALSE` to quiet this message.

```{r}
read_csv("a,b,c
1,2,3
4,5,6")
#> Rows: 2 Columns: 3
#> ── Column specification ────────────────────────────────────────────────────────
#> Delimiter: ","
#> dbl (3): a, b, c
#> 
#> ℹ Use `spec()` to retrieve the full column specification for this data.
#> ℹ Specify the column types or set `show_col_types = FALSE` to quiet this message.
#> # A tibble: 2 × 3
#>       a     b     c
#>   <dbl> <dbl> <dbl>
#> 1     1     2     3
#> 2     4     5     6



```


#Tidying data using Tidyr

```{r}
library(tidyverse)
```

```{r}
table1
#> # A tibble: 6 × 4
#>   country      year  cases population
#>   <chr>       <int>  <int>      <int>
#> 1 Afghanistan  1999    745   19987071
#> 2 Afghanistan  2000   2666   20595360
#> 3 Brazil       1999  37737  172006362
#> 4 Brazil       2000  80488  174504898
#> 5 China        1999 212258 1272915272
#> 6 China        2000 213766 1280428583
table2
#> # A tibble: 12 × 4
#>   country      year type           count
#>   <chr>       <int> <chr>          <int>
#> 1 Afghanistan  1999 cases            745
#> 2 Afghanistan  1999 population  19987071
#> 3 Afghanistan  2000 cases           2666
#> 4 Afghanistan  2000 population  20595360
#> 5 Brazil       1999 cases          37737
#> 6 Brazil       1999 population 172006362
#> # ... with 6 more rows
table3
#> # A tibble: 6 × 3
#>   country      year rate             
#> * <chr>       <int> <chr>            
#> 1 Afghanistan  1999 745/19987071     
#> 2 Afghanistan  2000 2666/20595360    
#> 3 Brazil       1999 37737/172006362  
#> 4 Brazil       2000 80488/174504898  
#> 5 China        1999 212258/1272915272
#> 6 China        2000 213766/1280428583

# Spread across two tibbles
table4a  # cases
#> # A tibble: 3 × 3
#>   country     `1999` `2000`
#> * <chr>        <int>  <int>
#> 1 Afghanistan    745   2666
#> 2 Brazil       37737  80488
#> 3 China       212258 213766
table4b  # population
#> # A tibble: 3 × 3
#>   country         `1999`     `2000`
#> * <chr>            <int>      <int>
#> 1 Afghanistan   19987071   20595360
#> 2 Brazil       172006362  174504898
#> 3 China       1272915272 1280428583


```


```{r}
# Compute rate per 10,000
table1 %>% 
  mutate(rate = cases / population * 10000)
#> # A tibble: 6 × 5
#>   country      year  cases population  rate
#>   <chr>       <int>  <int>      <int> <dbl>
#> 1 Afghanistan  1999    745   19987071 0.373
#> 2 Afghanistan  2000   2666   20595360 1.29 
#> 3 Brazil       1999  37737  172006362 2.19 
#> 4 Brazil       2000  80488  174504898 5.61 
#> 5 China        1999 212258 1272915272 1.67 
#> 6 China        2000 213766 1280428583 1.67

# Compute cases per year
table1 %>% 
  count(year, wt = cases)
#> # A tibble: 2 × 2
#>    year      n
#>   <int>  <int>
#> 1  1999 250740
#> 2  2000 296920

# Visualise changes over time
library(ggplot2)
ggplot(table1, aes(year, cases)) + 
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))


```


#Workshop 5 - Spatial data in R 

#Using R as GIS 

```{r}
# install and load your packages
install.packages("sf") 
install.packages("terra")
install.packages("tmap")


#load into R library
library(tidyverse)
library(sf) # simple features
library (terra) # for raster
library(tmap) # Thematic maps are geographical maps in which spatial data distributions are visualized
```

```{r}
#load the copepod data into R studio
library(readr)
dat <- read_csv("data-for-course/copepods_raw.csv")
dat
```


# A tibble: 5,313 × 11
   silk_id segment_no latitude longitude sample_time_utc  project route vessel  
     <dbl>      <dbl>    <dbl>     <dbl> <chr>            <chr>   <chr> <chr>   
 1       1          1    -28.3      154. 26/06/2009 22:08 AusCPR  BRSY  ANL Win...
 2       1          5    -28.7      154. 26/06/2009 23:12 AusCPR  BRSY  ANL Win...
 3       1          9    -29.0      154. 27/06/2009 0:17  AusCPR  BRSY  ANL Win...
 4       1         13    -29.3      154. 27/06/2009 1:22  AusCPR  BRSY  ANL Win...
 5       1         17    -29.7      154. 27/06/2009 2:26  AusCPR  BRSY  ANL Win...
 6       1         18    -29.8      154. 27/06/2009 2:43  AusCPR  BRSY  ANL Win...
 7       1         26    -30.4      153. 27/06/2009 4:52  AusCPR  BRSY  ANL Win...
 8       1         30    -30.7      153. 27/06/2009 5:57  AusCPR  BRSY  ANL Win...
 9       1         33    -31.0      153. 27/06/2009 6:45  AusCPR  BRSY  ANL Win...
10       1         37    -31.3      153. 27/06/2009 7:50  AusCPR  BRSY  ANL Win...
# ... with 5,303 more rows, and 3 more variables: meanlong <dbl>, region <chr>,
#   richness_raw <dbl>


```{r}
qfish_data <- read_csv("C:\\Users\\Rhea Arya\\Desktop\\QFISH.csv")
QFISH
qfish_data$Year_No <- as.integer(qfish_data$`Calendar Year`)
```

```{r}
library(ggplot2)
ggplot(dat) + 
  aes(x = longitude, y = latitude, color = richness_raw) +
  geom_point()


```

```{r}
ggplot(dat, aes(x = latitude, y = richness_raw)) + 
  stat_smooth() + 
  geom_point()
```


```{r}
sdat <- st_as_sf(dat, coords = c("longitude", "latitude"), 
                 crs = 4326)
```
```{r}
crs4326 <- st_crs(4326)
crs4326 # look at the whole CRS
crs4326$Name # pull out just the name of the crs
[1] "WGS 84"


```
```{r}
sdat
plot(sdat["richness_raw"])


```

```{r}
#using tmap

tm_shape(sdat) + 
  tm_dots(col = "richness_raw")


```


```{r}

```









